### 2018-03-12 manuscript_rearranged.tex

\documentclass[11pt]{article}

%==============Packages & Commands==============
\usepackage{graphicx}
\usepackage{fancyvrb}
\usepackage{tikz}
%%%<
\usepackage{verbatim}
%\usepackage[active,tightpage]{preview}
%\PreviewEnvironment{tikzpicture}
%\setlength\PreviewBorder{5pt}%

\usepackage{geometry}                		% See geometry.pdf to learn the layout options. There are lots.
% \geometry{a4paper}                   		% ... or a4paper or a5paper or ...
%\geometry{landscape}                		% Activat\usetikzlibrary{arrows}e for for rotated page geometry
%\usepackage[parfill]{parskip}    		% Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}				% Use pdf, png, jpg, or eps§ with pdflatex; use eps in DVI mode
								% TeX will automatically convert eps --> pdf in pdflatex
\usepackage{amssymb}

\usepackage[ruled,vlined]{algorithm2e}
\usetikzlibrary{arrows}
\usepackage{alltt}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{indentfirst}
\usepackage[longnamesfirst]{natbib} % For references
\bibpunct{(}{)}{;}{a}{}{,} % Reference punctuation
\usepackage{changepage}
\usepackage{setspace}
\usepackage{booktabs} % For tables
\usepackage{rotating} % For sideways tables/figures
\usepackage{amsmath}
\usepackage{multirow}
\usepackage{color}
\usepackage{dcolumn}
\usepackage{comment}
%\usepackage{fullwidth}
\newcolumntype{d}[1]{D{.}{\cdot}{#1}}
\newcolumntype{.}{D{.}{.}{-1}}
\newcolumntype{3}{D{.}{.}{3}}
\newcolumntype{4}{D{.}{.}{4}}
\newcolumntype{5}{D{.}{.}{5}}
\usepackage{float}
\usepackage[hyphens]{url}
%\usepackage[margin = 1.25in]{geometry}
%\usepackage[nolists,figuresfirst]{endfloat} % Figures and tables at the end
\usepackage{subfig}
\captionsetup[subfloat]{position = top, font = normalsize} % For sub-figure captions
\usepackage{fancyhdr}
%\makeatletter
%\def\url@leostyle{%
%  \@ifundefined{selectfont}{\def\UrlFont{\sf}}{\def\UrlFont{\small\ttfamily}}}
%\makeatother
%% Now actually use the newly defined style.
\urlstyle{same}
\usepackage{times}
% \usepackage{mathptmx}
%\usepackage[colorlinks = true,
%						bookmarksopen = true,
%						pagebackref = true,
%						linkcolor = black,
%						citecolor = black,
% 					urlcolor = black]{hyperref}
%\usepackage[all]{hypcap}
%\urlstyle{same}
\newcommand{\fnote}[1]{\footnote{\normalsize{#1}}} % 12 pt, double spaced footnotes
\def\citeapos#1{\citeauthor{#1}'s (\citeyear{#1})}
\def\citeaposs#1{\citeauthor{#1}' (\citeyear{#1})}
\newcommand{\bm}[1]{\boldsymbol{#1}} %makes bold math symbols easier
\newcommand{\R}{\textsf{R}\space} %R in textsf font
\newcommand{\netinf}{\texttt{NetInf}\space} %R in textsf font
\newcommand{\iid}{i.i.d} %shorthand for iid
\newcommand{\cites}{{\bf \textcolor{red}{CITES}}} %shorthand for iid
%\usepackage[compact]{titlesec}
%\titlespacing{\section}{0pt}{*0}{*0}
%\titlespacing{\subsection}{0pt}{*0}{*0}
%\titlespacing{\subsubsection}{0pt}{*0}{*0}
%\setlength{\parskip}{0pt}
%\setlength{\parsep}{0pt}
%\setlength{\bibsep}{2pt}
%\renewcommand{\headrulewidth}{0pt}

%\renewcommand{\figureplace}{ % This places [Insert Table X here] and [Insert Figure Y here] in the text
%\begin{center}
%[Insert \figurename~\thepostfig\ here]
%\end{center}}
%\renewcommand{\tableplace}{%
%\begin{center}
%[Insert \tablename~\theposttbl\ here]
%\end{center}}

\newcommand\independent{\protect\mathpalette{\protect\independenT}{\perp}}
\def\independenT#1#2{\mathrel{\rlap{$#1#2$}\mkern2mu{#1#2}}}
\newcommand{\N}{\mathcal{N}}
\newcommand{\Y}{\bm{\mathcal{Y}}}
\newcommand{\bZ}{\bm{Z}}

\usepackage[colorlinks = TRUE, urlcolor = black, linkcolor = black, citecolor = black, pdfstartview = FitV]{hyperref}

%============Article Title, Authors==================
\title{\vspace{-2cm} Inference on the Effects of Observed Features
\\ in Latent Space Models for Networks }

\author{ Zachary Jones \and Matthew Denny \and Bruce Desmarais \and Hanna Wallach} \date{\today}

%===================Startup=======================
\begin{document}
\maketitle

%=============Abstract & Keywords==================

\begin{abstract}

\noindent Due to the complex interdependence found in network data,
networks scholars draw upon an increasingly sophisticated
toolkit for building models of networks. The Latent space model (LSM) for
network data combines a conventional regression model with an embedding of actors in a latent space. In the LSM, the expected relationship between two actors is dependent on observed covariates and a function of their positions in the latent space. In numerous applications, researchers have assumed
that the latent spatial embedding can control for latent
confounding structure. latent network structure includes unmeasured actor attributes that can be represented spatially, as well as network dependencies such as reciprocity and transitivity. There has been little research that considers the LSM's
performance in adjusting for latent network structure when making inferences about covariates. We investigate the
LSM's performance via a simulation study. In the presence of an
unmeasured actor attribute that can be modeled perfectly using a latent space,
we find that the LSM exhibits unacceptable levels of bias and Type 1 error for inferences on measured covariates. However, the prediction error of the LSM when latent network structure is present is substantially lower in most cases. We
conclude that the LSM is most appropriate for exploratory or predictive modeling, unless it can be assumed that all potential confounders have been measured.
%\footnote{This work was supported by National Science Foundation
%grants DGE-1144860, SES-1558661, SES-1619644, and CISE-1320219. Any
%opinions, findings, and conclusions or recommendations are those of
%the authors and do not necessarily reflect those of the sponsor.}
\end{abstract}
\thispagestyle{empty}
% \doublespacing
% Description of the possible challenges
\section{Introduction}

Statistical models in which the outcome data (i.e., the dependent variable) consists of network data have proliferated, and grown increasingly
sophisticated, over the past decade. The defining feature of network data that renders conventional statistical models---namely, regression models---inappropriate, is that ties in a network exhibit complex forms of interdependence such as reciprocity, transitivity, and
homophily. If not accounted for, these complex dependencies can lead to biased estimates and
errors in hypothesis testing, much in the way that omitted variable
bias can affect results in conventional regression models
\citep{ward2007disputes,kinne2014,cranmerisq,hays2010spatial}.  We refer to network dependencies that cannot be modeled with observed covariates as {\em latent network structure}. A
number of statistical modeling frameworks have been proposed to
account for latent network structure in network data. These include, but are not limited to, the exponential random
graph model (ERGM) {\bf [UPDATE STRING CITES TO LOOK OUTSIDE POLISCI]} \citep[e.g.,
][]{lazer2010,cranmer2011pa,desmarais2012psj}, the latent space model
(LSM) \citep[e.g.,
][]{ward2007disputes,ward2007persistent,kirkland2012multimember}, and
the stochastic actor oriented model (SAOM) \citep[e.g.,
][]{berardo2010ajps,kinne2014}.

The approach to adjusting for latent network structure in the ERGM and SAOM is
quite similar to adjusting for confounding covariates in regression
modeling. The researcher specifies a set of dependencies that (s)he
hypothesizes to be important in the generative model for the
network. These dependencies are then explicitly included in a model
that simultaneously represents the effects of observed covariates
\citep{cranmer2011pa}. The LSM takes a different approach, which
involves the incorporation of latent variables to model latent network structure. The LSM has an advantage over ERGM and SAOM in that
researchers need not develop a precise set of hypothesized dependencies in
order to model latent network structure. However, this advantage hinges upon the capacity for the
LSM to differentiate the effects of latent network structure 
from those of the observed covariates.

Despite their growing popularity, few studies exist that investigate
the performance of these models in adjusting for confounding network
structure. In the current study, we focus on the LSM, examining its
performance in reducing estimation and inferential errors regarding
the effects of observed covariates via adjustment for confounding
network structure. There are two reasons that using the LSM may lead
to increased error---one may result in Type 1 inferential error, and
the other in Type 2 error. First, if the latent
network structure is truly correlated with the observed covariates,
the unobserved structure that can be correlated with the observed
variable may be attributed to the observed variable. Under this
condition the latent space parameters are used to model sources of
variation that cannot be attributed to the observed covariate, and
inferences regarding the observed covariates remain subject to omitted
variable bias, which leads to an inflated Type 1 inferential error
rate. Second, the latent configurations inferred may result in a
representation of the network wherein a node's position in the latent
space is spuriously correlated with the observed covariates (i.e., the latent space crowds out the effects of observed covariates). Inferring
latent variables that are correlated with observed covariates would
lead to reduced efficiency in estimating observed covariate effects,
and result in a high Type 2 error rate.

%HERE, ADD OVERVIEW OF OUR PAPER

% CUT: not sure where to put this
%Despite the power of the LSM in representing network structure that is
%not modeled by the observed covariates, and its performance in
%predictive ties predicting ties, we have reason to doubt the LSM's
%capacity to adjust for confounding network structure.

\section{The Latent Space Model}

The LSM, introduced by \citet{hoff2002latent}, is used to estimate the
effect of covariates in the presence of latent network structure. Here
the distance function $|z_i - z_j|$ represents latent network
structure as homophily with respect to latent variables.  The distance
function is additively combined with a regression on observed dyadic
covariates, $x_{ij}$, to form a linear predictor for tie prediction.
As with a GLM, a link function, $g^{-1}$, maps this linear predictor
to the appropriate edge distribution, giving

$$\mathbb{E}(y_{ij} | x_{ij}) = g^{-1}(\alpha + \beta x_{ij} - |z_i - z_j|).$$

This ``Euclidean'' LSM is the original form proposed by
\citet{hoff2002latent}, and, as far as we can tell, the most commonly
used specification in the literature.

% CUT: what's the point of this paragraph?
%However, the LSM has been extended in several important
%directions. The latent space has been represented as a $k$-dimensional
%Euclildean space and by latent factors with a multiplicative metric
%\citep{hoff2002latent, hoff2004modeling, hoff_2005_jasa,
%  hoff2009multiplicative}. Additional structure has been introduced by
%adding random effects, which, for example, may involve sender or
%receiver specific effects for directed networks which capture
%differential activity rates amongst nodes
%\citep{hoff2003random}. Within this framework \cite{westveld2011mixed}
%models dynamic network data by treating the latent space as a
%stochastic process. \cite{handcock2007model} enable the LSM to model
%clustering that is not representable as homophily (i.e., stochastic
%equivalence) by combining the LSM with latent cluster
%models. \cite{hoff2008modeling} shows that the LSM and latent cluster
%models are special cases of an ``eigenmodel.'' That is, an
%eigendecomposition of a symmetric sociomatrix can be used to represent
%both latent space and cluster models, but not
%vice-versa. \citet{hoff2015dyadic} consolidates, extends, and
%implements several of these extensions in the \texttt{amen} package
%\citep{amen} for \R \citep{minhas2016inferential, hoff2015dyadic}.

The LSM has seen use in a variety of fields in which network data is
common, particularly the social sciences. The apparent appeal of the
LSM appears to be driven primarily by the LSM's usefulness in modeling
transitivity (i.e., clustering) and homophily, which are ubiquitous in
social networks. In political science the LSM and variations on the
form developed in \cite{hoff2002latent} have been used to estimate the
effect of democracy on the probability of a militarized interstate
dispute \citep{ward2007disputes}, the amount of portfolio investment
between states \citep{cao2013democracies}, and the effect of
multimember districts on the probability of collaboration between
state legislators in the United States
\citep{kirkland2012multimember}. Variations of the LSM developed for
networks measured over time have been applied to the study of
international trade, wherein the effects of various features of
trading partners are estimated \citep{ward2013gravity}. In ecology the
LSM has been used to study the sociality of elephants
\citep{vance2009social} and orcas \citep{fearnbach2014spatial}, birds
\citep{nomano2015unrelated}, to discover ecological communities
\citep{fletcher2011social, fletcher2013network}, and to study food
webs \citep{chiu2011unifying}. In epidemeology it has been used to
identify clusters of infected persons for later isolation
\citep{zhang2015cluster} and to study patterns of interaction amongst
physicians \citep{paul2014results}. In marketing and business research
it has been used to study inter-group trust \citep{dass2011impact},
optimal bundling and pricing of goods and brands for retailers
\citep{dass2012assessing}. It has been used to describe topic-specific
patterns of interaction in e-mail communication networks
\citep{krafft2012topic}.  It has been used to estimate the effects of
education policy interventions on the structure of friendship networks
among students \citep{sweet2011modeling}. Lastly, in neuroscience it
has been proposed as a method for modelling fMRI data
\citep{simpson2013analyzing}.

%Below we review several studies in which researchers have relied on
%the latent space modeling framework to adjust for unmeasured
%confounding structure.

Although the LSM has often been used as an exploratory or predictive
model, it has been applied in many cases for explanatory causal
modeling---to reduce estimation and/or inferential error with respect
to the effects of observed covariates. We do not claim that the
methodological literature in which the LSM was introduced or extended
has presented the LSM as capable of adjusting for unmeasured
confounding structure. However, as we document below, many researchers
who have applied the LSM claim that it has the capacity to adjust for
confounders, and use it for this purpose. For example
\citet{ward2007disputes}, in a high profile example, argue that the
the LSM improves inference about the effects of democracy,
international trade, and participation in international organizations
on the probability of inter-state conflict.

\begin{quote}
``The history of international disputes, and consequently the extant
  data on militarized interstate disputes, is replete with \ldots
  dependencies. We formally incorporate and estimate the extent of
  these \ldots dependencies in our model of the Kantian peace in order
  to more precisely determine the effects of the Kantian tripod on
  international conflict'' \citep[p. 585]{ward2007disputes}
\end{quote}

Likewise, similar claims are made in \citet{dorff2013networks,
  cao2013democracies, ward2013gravity, kirkland2012multimember}, and
\citet{cao2016transnational}.

\begin{quote}
``For the most part, however, most dyadic research in international
  relations ignores the essential features of dyads in that they fail
  to satisfy the assumption of independence or, by construction, have
  missing data but ignore its effects: both of these bias the results
  in a fundamental way'' \citep[p. 2]{dorff2013networks}.
\end{quote}

\begin{quote}
``This approach combines a network analysis with a standard-looking
  regression to permit us to access the importance of our explanatory
  factors without having them biased by the interdependencies in the
  network we are studying'' \citep[p. 15]{cao2013democracies}.
\end{quote}

\begin{quote}
``The presence of \ldots dependence implies misspecification and a
  high likelihood of bias in most current applications. Building on
  the latent space framework we model the world trading system without
  assuming a particular network structure or the sufficiency of
  particular network statistics'' \citep[p. 20]{ward2013gravity}.
\end{quote}

\begin{quote}
``\ldots the latent space model allows for the assessment of distance
  between two unconnected actors while simultaneously controlling for
  the interdependence inherent in network data. This interdependence
  in latent space positions allows the model to control for common
  network effects like reciprocity or transitivity that would
  ordinarily bias results'' \citep[p. 336]{kirkland2012multimember}.
\end{quote}

\begin{quote}
``Ignoring third-order dependence in dyadic data and treating dyads
  Germany-France, France-Italy, and Germany-Italy as independent
  observations can cause bias in parameter estimates (Hoff 2005). The
  statistical literature has proposed a series of latent space to
  control for autocorrelation among dyadic observations (Hoff et
  al. 2002; Hoff 2005). Countries? unobserved characteristics are
  captured by latent vectors...'' \citep[p. 17]{cao2016transnational}.
\end{quote}

The above examples are drawn from political science, but the results
of our analysis have implications in other fields. We see a similar
logic for using the LSM articulated in recent work in ecology by
\citet[p. 989]{nomano2015unrelated}.

\begin{quote}
``\ldots can create artificially exaggerated synchrony rates
  regardless of motivation for signaling, which can be modeled with a
  term known as ``transitivity'' in the social network literature. A
  latent space model (Krivitsky et al. 2009) was used to examine the
  propensities for synchrony between helper males and the primary male
  while accounting for the variability deriving from the
  transitivity.''
\end{quote}

And in sociology \citet[p. 157]{spillane2017elephant},

 \begin{quote} ``the models include latent space positions that implicitly control for structural aspects of networks, such as reciprocity and centrality, allowing for estimation of covariate effects (Hoff, Raftery, and Handcock 2002).''
 \end{quote}

 and  \citet[p. 107]{berlusconi2017determinants}

 \begin{quote} ``Higher order dependencies can be taken into account by assigning each country a position in a ?social space,? based on its (unobserved) characteristics and the presence of other ties. Conditional independence between observations can then subsequently be assumed, given the countries? latent position within the social space (Hoff et al., 2002).''
 \end{quote}

Though this list of quotations is by no means complete, the last
example to which we point comes from the business literature, and uses
the bilinear form of the latent space model
\citep[p. 7]{dass2011impact}.

\begin{quote}
``This model controls for higher order team dynamics using the
  bilinear component $z_i'z_j$ and also allows both trustor
  ($x_{tor,i}$ ) and trusted ($x_{ted,j}$ ) characteristics to be
  investigated along with the dyadic covariates ($x_{d,i,j}$).''
\end{quote}

The latent space framework offers an attractive general purpose approach to adjusting for confounding structure in network models, as, unlike the major alternative framework for network modeling (ERGM), using the latent space model does not require the researcher to specify a set of network dependencies for which to control.  However, the latent variables in the LSM are free parameters that will be inferred to explain variation that is not explained by the observed covariates. As such, the LSM framework is not designed to encourage latent network structure to displace observed covariate effects when the latent structure includes confounders. As we note above, recent work on latent space approaches for network modeling describes exactly this modeling characteristic---that the latent variables are used to explain variation that cannot be explained by observed covariates. In a simulation study that follows, we examine whether the LSM can be used to adjust for confounding in a simple and ideal scenario..

\section{Simulation Study}

We use a simulation study to evaluate the LSM's performance at
inferring covariate parameters in the situation in which we have some
observed covariates as well as omitted network structure which can be
represented using a Euclidean latent space. We are particularly
interested in whether the LSM can adjust for confounding when the
confounding variable is unmeasured. In the simulation design, we study
how the level of confounding affects the performance of the LSM
relative to the GLM. We evaluate performance in terms of bias and
inferential error. Since the LSM has also been used for predictive
model selection, we also evaluate predictive performance
\citep{ward2013gravity, fletcher2011social, fletcher2013network,
  chiu2011unifying}.

\subsection{Simulation Design}

+For simplicity and computational efficiency we consider a unidimensional latent space, that is, we create a single unmeasured variable that can be represented using the Euclidean parameterization of the LSM. We expect our results to generalize to higher dimensions and other variants of the LSM since each variant involves the use of unconstrained latent variables to represent unmeasured network structure. To generate an observed covariate that exhibits a controllable degree of colinearity with the latent network structure we follow a four-step process. First, we simulate unidimensional positions for each node, drawing from a standard normal distribution, and calculate the Euclidean distance $\mathbf{d}$ between each pair of positions. Second, we generate a target covariance matrix ($\Sigma$) using the $c$-vine method of \cite{lewandowski2009generating} which generates a uniform draw from the space of positive-definite matrices. The $c$-vine method makes use of a parameter $\eta$, which when equal to $1$ generates draws from uniformly from the space of positive-definite matrices, making the correlation between the latent positions and the measured covariate 0 in expectation. As $\eta$ increases above $1$ draws of $\Sigma$ have increasingly weak correlation\footnote{We show the distribution of correlations under each value of $\eta$ used in our simulation in Figure \ref{fig:vine}.}. Third, given the target covariance matrix ($\Sigma$) among the covariates and distances, $\langle \mathbf{x}, \mathbf{d} \rangle$, we derive the conditional mean vector and covariance, assuming that $\mathbf{x}$ has a normal distribution given $\mathbf{d}$ (see \cite[pp. 116--117]{eaton1983} for the conditional normal derivation).
+Fourth, we simulate $x$ as a normal random variable with the respective conditional means and covariance. Finally we standardize $x$ to have zero mean and unit variance to ensure that the scale of estimation errors is comparable across simulations.

We consider three exponential family distributions from which the adjacency matrix entries are drawn: Gaussian, Bernoilli, and Poisson. We also vary the size of the network under study, considering networks with $n = 25, 50,$ and $100$ nodes. To vary the degree of confounding attributable to the latent positions, we consider three values of the collinearity parameter $\eta$: $1$, $100$, and $1,000,000$. Here  where $\eta = 1$ gives a matrix where the correlation is in expectation $.5$, $\eta = 100$ results in moderate collinearity between the observed covariate and the latent distances, and $\eta = 1,000,000$ corresponds to independence between the observed covariate and the latent distances. See Figure \ref{fig:vine} for the distribution of the absolute value of the correlation generated at each value of $\eta$.

\begin{figure}
\includegraphics[width=\textwidth]{figures/max_r_vine.png}
\caption{The distribution of the absolute value of the correlation between the observed covariate and the latent distances. \label{fig:vine}}
\end{figure}

We consider the LSM with several different priors on the coefficient
for the observed covariate and the latent space. We set the prior
variance of $\beta$ to be either $1$ or $10$ and use a diffuse normal
prior on the latent space.

The LSM is estimated using the canonical implementation in the
\texttt{latentnet} package in \texttt{R} \citep{latentnet}. For each
iteration of the simulation, we run 10,000 burn in iterations,
followed by 1,000,000 iterations of the sampler. Every 100th post burn
in iteration is saved. Convergence in the log probability of the model
is assessed using the Geweke diagnostic in the \texttt{coda} package
\citep{coda, geweke1991evaluating}. If the convergence criterion is
satisfied, the simulation continues to the next set of arguments,
otherwise the number of iterations is doubled. If the convergence
criterion is still not satisfied, then the aforementioned step in the
simulation is flagged for review. However, this was not necessary in
any cases. At each point in the simulation's parameter space, we
execute 1,000 Monte Carlo iterations. \footnote{One of the primary
  difficulties in executing the above simulation design is the
  computational cost of estimating the LSM. We utilize the
  \texttt{BatchExperiments} \R package to construct and execute our
  computational experiments on a Torque cluster
  \cite{bischl2015batchjobs}.}

We compute the MLE estimated by iteratively reweighted least squares
via the \texttt{glm} function in \texttt{R} \citep{rcore}. For the LSM
we use the posterior mode as our point estimate. In the cases where
edges are Bernoulli and there is omitted network structure, we scale
the estimates using the reciprocal of the bias of $\beta$ when
$\mathbf{x}$ and $\mathbf{d}$ are independent: $\sqrt{\frac{3.28 +
    \beta^2 \text{Var}(\mathbf{d})}{3.29}}$, where $3.29$ is the
variance of a standard logistic distribution. We do this to adjust for
bias in the coefficient estimate that arises due to the lack of a
scale coefficient in logistic regression (See the derivation of the
bias under an independent but omitted covariate in
\cite{mood2010logistic}).

For each combination of simulation conditions we evaluate the mean
square prediction error of the model on new edges drawn condtional on
$\mathbf{x}$ and $\mathbf{d}$: $\frac{1}{n}\sum_{i=1}^{n}(y_{i} - \hat{y}_i)^2$, the bias of the estimated coefficient
for the measured covariate $\frac{1}{n} \sum_{i=1}^n (\beta - \hat{beta})$, and the Type-1 and 2 error
rates. These error rates are estimated by computing coverage: $\frac{1}{m} \sum_{i=1}^m\mathbb{I}(\beta \in \text{HPD}(\hat{\beta}_i))$, where $\text{HPD}(\hat{beta})$ is the region of highest posterior density which covers $95\%$ of the marginal posterior distribution and $m$ is the number of samples from said posterior \cite{turkkan1993computation}. When in the simulation $\beta = 0$, $1 - P(\text{coverage})$ gives the type-1 error rate, and $P(\text{coverage})$ givesthe type-2 error rate when $\beta = 1$. Hence we consider $\beta$ to be statistically
significant at the 0.05 level if the 95\% credible interval excludes 0.

Recent work on the LSM has explicitly addressed the potential problem
of Type 2 error. The AMEN framework for latent variable modeling of
networks is explicitly designed to model structure in the
residuals---structure that is leftover after accounting for the
observed covariates. \citet[p. 43]{hoff2015dyadic} notes that the
latent factors represent patterns in the network that, ``aren't
explained by the known regressors.''
\citet[pp. 12--13]{minhas2016inferential} also describe how AMEN is
designed such that the multiplicative effects, ``capture higher-order
dependence patterns that are left over [in the stochastic linear
  predictor] after accounting for any known covariate information.''
The approach incorporated in AMEN is effective at avoiding correlation
between the observed covariates and latent factors, which would result
in Type 2 error. However, there have been no methodological
innovations to avoid Type 1 error in latent variable modeling with
networks. Through a simulation study we find that the LSM, as
implemented in both the R \citep{R} packages \texttt{latentnet}
\citep{latentnet} and \texttt{amen} \citep{amen}, performs poorly in
terms of Type 1 error when the unmeasured network structure confounds
the relationship between the observed covariate and the network.

We also repeat this simulation study using \texttt{amen}
\citep{hoff2015dyadic, minhas2016inferential}. We omit the Poisson
family, and, due to the computational cost of the simulation, we
consider only networks with 25 nodes or 100 nodes. Rather than
constructing an omitted variable as described above, we draw from a
standard normal distribution $\mathbf{u}$, and, because we are
considering symmetric networks, take the outer product of $\mathbf{u}$
with itself to construct the latent factor. We take 100,000 MCMC
samples after 500 burn-in iterations, and we assess convergence using
the diagnostic of \cite{raftery1992practical}.

\subsection{Results}

To evaluate estimation error we compute the bias. Figures
\ref{fig:estimation_ls} and \ref{fig:estimation_nls} shows the
results. In Figure \ref{fig:estimation_ls} we can see that with a
uniform distribution over the correlation between the distances and
the observed covariate, all methods (except the true model wherein the
distances are included via an observed covariate) show substantial
bias. In most cases the LSM performs better than the GLM, sometimes by
as much as $50\%$. However, in absolute terms, the amount of bias is
large. Our results demonstrate that using the LSM will not permit the
discovery of and adjustment for the true latent positions when the
observed covariates and latent distances are confounded.  For both the
GLM and the LSM, bias decays rapidly as the degree of correlation
between the distances and the measured covariate decreases. In
\ref{fig:estimation_nls} we can see that when there is no omitted
network structure, the LSM exhibits bias greater than that of the GLM.

\begin{figure}
\includegraphics[width=\textwidth]{figures/estimation_ls.png}
\caption{The bias of estimates of the effect of the observed covariate $\mathbf{x}$ when there is an ommitted variable. The $x$-axis gives the value of the parameter $\eta$ which controls the degree of dependence between $\mathbf{x}$ and ommitted covariate. Lower values of $\eta$ indicate higher levels of dependence between the observed and ommitted covariate. The $y$-axis gives a Monte Carlo estimate of the bias. The number of nodes are indicated in the top panels, while the distributional family of the edges is shown on the right panel. Each panel represents 4 values of $\eta$ with 1,000 Monte Carlo iterations executed at each point.
\label{fig:estimation_ls}}
\end{figure}

\begin{figure}
\includegraphics[width=\textwidth]{figures/estimation_nls.png}
\caption{The bias of estimates of the effect of the observed covariate $\mathbf{x}$ when there is no ommitted variable.
\label{fig:estimation_nls}}
\end{figure}

Figures \ref{fig:inference_type_1} and \ref{fig:inference_type_2} show the inferential error rates of the LSM with different priors for $\beta$, as well as a GLM. Type-1 error rates shown in Figure \ref{fig:inference_type_1} for the LSM are in general somewhat lower than those of a GLM when an unmeasured covariate exists. Under uniform correlation between the omitted network structure and the observed covariate, the Type-1 error rate is often more than 10 times greater than the nominal rate of 0.05. This illustrates that the LSM cannot be considered a viable substitute to measuring the omitted structure, as the true model exhibits a Type-1 error rate that matches the nominal rate. Type-2 error rates, shown in Figure \ref{fig:inference_type_2} are also substantially above their nominal rate when there is a uniform distribution on the strength of confounding, though the inflation is not as bad as that of the Type-1 rates. These results indicate that, in the presence of unobserved confounders that can be represented through the LSM, the LSM does not adequately correct the inferential errors that would arise due to omitted structure among the dyads in the network.

\begin{figure}
\includegraphics[width=\textwidth]{figures/inference_type_1.png}
\caption{Monte Carlo estimates of the Type-1 error regarding the effect of the observed covariate $\mathbf{x}$ are shown on the $y$-axis. Here, $\beta = 0$ and the error rate shown in each panel in that row gives 1 minus the probability of a 95\% confidence region (for the LSM) or interval (for the GLM) including $0$, giving the probability that a true null hypothesis of $\beta = 0$ is falsely rejected. \label{fig:inference_type_1}}
\end{figure}

\begin{figure}
\includegraphics[width=\textwidth]{figures/inference_type_2.png}
\caption{Monte Carlo estimates of the Type-2 error regarding the effect of the observed covariate $\mathbf{x}$ are shown on the $y-axis$. Here, $\beta = 1$, and the error rate shown in each panel in each row gives the probability of the probability/confidence intervals covering $0$, giving the probability of accepting a false null hypothesis. \label{fig:inference_type_2}}
\end{figure}

The last property we consider in comparing the GLM and LSM is generalization error---the performance exhibited by the models in predicting new data that was drawn from the same model that generated the data used for estimation (i.e., out of sample predictive performance). We estimate generalization error as the expected prediction error on new edges generated using a fixed set of latent positions for the nodes and a fixed observed covariate. Generalization error results are shown in Figure \ref{fig:generalization}. In nearly all cases the LSM substantially outperforms the GLM, especially when the omitted covariate is not highly collinear with the observed covariate. In nearly all cases it performs as well as the true model.

\begin{figure}
\includegraphics[width=\textwidth]{figures/generalization.png}
\caption{The $x$-axis shows a Monte Carlo estimate of the mean square error for edge values drawn from the appropriate distribution with the observed covariate $\mathbf{x}$ and the latent positions $\mathbf{z}$ fixed. In the Bernoulli case the Brier score is computed and in the Poisson and normal cases the mean-square-error. \label{fig:generalization}}
\end{figure}

In our repetition of this simulation with \texttt{amen} we find similar results, with the exception of type-2 error rates, which, as expected, are substantially lower for \texttt{amen}. However, type-1 error rates remain elevated, as is bias in the case where there is a latent dependence structure that is correlated with the observed covariate. \texttt{amen} does perform better in the case where it is used despite the absence of such a latent dependence structure though, relative to \texttt{latentnet}, which is to be expected given that it outperforms \texttt{latentnet} in terms of type-2 error. Like the LSM estimated with \texttt{latentnet}, \texttt{amen} performs well in terms of generalization error. The results of our simulations with \texttt{amen} are available in the Online Appendix.

\section{Conclusion}

Based on our simulation study, we conclude that the primary advantage of using the LSM---relative to a dyadic regression model in which the network is fit using observed covariates only---is that the latent space provides an efficient complement to observed covariates when it comes to fitting and predicting the network; \texttt{amen} increases this advantage.
Considering both (1) that the LSM does not substantially reduce bias or inferential error relative to the GLM, and (2) that the LSM exhibits strong predictive performance, we conclude that the latent positions are used to explain the systematic variation that cannot be explained through the observed covariates. This is of considerable use in research focused on developing a predictive model, or fitting and exploring latent positions. However, since the latent positions are inferred to complement the observed covariates---explaining residual variation---the LSM is ill-suited to adjust for confounding network structure, as adjusting for confounding would require that the latent positions explain variation that could otherwise be attributed to the observed covariates.

The key recommendation arising from our results is that researchers not use the LSM as an alternative to measuring potential confounding variables and/or explicitly modeling the endogenous network structure. The LSM cannot control for unmeasured confounders. This property is, perhaps, not surprising. Given that all priors are centered at zero, the parameter regions that exhibit high posterior probability will be those that explain the observed network (i.e., high likelihood) using minimal departures from zero (i.e., high prior probability). Since explaining the observed network using a covariate requires moving one parameter value away from zero, whereas explaining the network using a latent dimension requires moving many latent coordinates away from zero, the LSM exhibits an inherent preference to explain network structure using observed covariates over latent coordinates. One way to address this issue would be to use a highly diffuse prior for the latent positions, but as we discuss above, that is not possible with the LSM. If the LSM is to be used as a tool for drawing inferences regarding observed covariate effects while adjusting for confounding network structure, further development is needed to enable the LSM to avoid bias and inferential error under confounding.

\newpage

\bibliographystyle{apsr}
\bibliography{ref}

\newpage
\section*{Online Appendix}
Below we replicate the simulation study results figures using the results from our simulation with the recently developed \texttt{amen} package. Note that these simulations exclude (1) the Poisson family, which is not available in the \texttt{amen} package, and (2) the 50-node networks, which we left out in the interest of lessening the computational burden of the simulation.

\begin{figure}
\includegraphics[width=\textwidth]{figures/estimation_ls_amen.png}
\caption{The bias of estimates of the effect of the observed covariate $\mathbf{x}$ when there is an ommitted variable. The $x$-axis gives the value of the parameter $\eta$ which controls the degree of dependence between $\mathbf{x}$ and ommitted covariate. Lower values of $\eta$ indicate higher levels of dependence between the observed and ommitted covariate. The $y$-axis gives a Monte Carlo estimate of the bias. The number of nodes are indicated in the top panels, while the distributional family of the edges is shown on the right panel. Each panel represents 4 values of $\eta$ with 1,000 Monte Carlo iterations executed at each point.
\label{fig:estimation_ls_amen}}
\end{figure}

\begin{figure}
\includegraphics[width=\textwidth]{figures/estimation_nls_amen.png}
\caption{The bias of estimates of the effect of the observed covariate $\mathbf{x}$ when there is no ommitted variable.
\label{fig:estimation_nls_amen}}
\end{figure}

\begin{figure}
\includegraphics[width=\textwidth]{figures/inference_type_1_amen.png}
\caption{Monte Carlo estimates of the Type-1 error regarding the effect of the observed covariate $\mathbf{x}$ are shown on the $y$-axis. Here, $\beta = 0$ and the error rate shown in each panel in that row gives 1 minus the probability of a 95\% confidence region (for the LSM) or interval (for the GLM) including $0$, giving the probability that a true null hypothesis of $\beta = 0$ is falsely rejected. \label{fig:inference_type_1_amen}}
\end{figure}

\begin{figure}
\includegraphics[width=\textwidth]{figures/inference_type_2_amen.png}
\caption{Monte Carlo estimates of the Type-2 error regarding the effect of the observed covariate $\mathbf{x}$ are shown on the $y-axis$. Here, $\beta = 1$, and the error rate shown in each panel in each row gives the probability of the probability/confidence intervals covering $0$, giving the probability of accepting a false null hypothesis. \label{fig:inference_type_2_amen}}
\end{figure}

\begin{figure}
\includegraphics[width=\textwidth]{figures/generalization_amen.png}
\caption{The $x$-axis shows a Monte Carlo estimate of the mean square error for edge values drawn from the appropriate distribution with the observed covariate $\mathbf{x}$ and the latent positions $\mathbf{z}$ fixed. In the Bernoulli case the Brier score is computed and in the normal case the mean-square-error. \label{fig:generalization_amen}}
\end{figure}

\end{document}